Formalizing Social Engineering attacks in the Symbolic Model

### 2023-07-23 Seminar/article.tex
Some retouches

Given a multiset $\Gamma_t = \{ F_0, ..., F_n \}$ and a sequence of multisets $trace_t = \langle a_0, ..., a_{t-1} \rangle$ at a time $t$, we can define a rewrite rule as a triple of multisets $RR = \langle L, A, R \rangle$ (written as $RR= L -[ A ] \rightarrow R$) such that:

\begin{itemize}
    \item we can apply $RR$ to $\Gamma_t$ if there is at least one ground instance (i.e. an instance with no variables) $rr = l -[ a ] \rightarrow r$ of $RR$ so that $l \subseteq^{\#} \Gamma_t$
    \item applying $rr$ to $\Gamma_t$ yields to a new state $\Gamma_{t+1}$ and an increased trace $trace_{t+1}$ obtained as
\begin{align*}
    \Gamma_{t+1} &= \Gamma_t \setminus^{\#} lin(l) \cup^{\#} r\\
    trace_{t+1} &= \langle a_0, ..., a_{t-1}, a \rangle
\end{align*}
\end{itemize}
in which $\setminus^{\#}$ and $\cup^{\#}$ are the multiset equivalent operations for set difference and union and $lin(l)$ is the multiset of linear facts belonging to $l$ (notice that persistent facts are never removed from the state). From now on, we will refer to $L$, $R$ and $A$ as the multisets of \textit{premises}, \textit{conclusions} and \textit{action facts} of a rule (in this order).

@@

\begin{definition}[Set of possible traces]
    G Given a set of labelled rewriting rules $P = \{(N_1,RR_1),$ $...,(N_m,RR_m)\}$, we define the set of possible traces generated by $P$ as
    \begin{align*}
        traces(P) = \{ &\langle A_1,...,A_n \rangle |\\
        & \exists S_1,...,S_n . \varnothing ^{\#} \xrightarrow[]{A_1} S_1 \xrightarrow[]{A_2} ... \xrightarrow[]{A_n} S_n\\
        & \textrm{and no ground istance of } Fresh()\\
        & \textrm{is used twice }\}
    \end{align*}
    where $A_i$ is $RR_i$'s action facts multiset.
\end{definition}

@@

To better understand how such trace (both observable and not) is derived from a protocol's execution, we can take a look at the following example rewriting system:
\begin{align*}
    P = \{ &(N_1, \varnothing^{\#}-[\{\{\textrm{Init}(0)\}\}] \rightarrow \{\{A(0)\}\}),\\
    &(N_2, \{\{A(x)\}\}-[\varnothing^{\#}] \rightarrow \{\{B(x)\}\})\\
    &(N_3, \{\{B(x)\}\}-[\{\{\textrm{Concl}(x)\}\}] \rightarrow \varnothing^{\#})\}
\end{align*}

Let us assume we are starting with an empty state $\Gamma_0$ and apply rules $N_1, N_2, N_1, N_3$, in this order. The state would evolve as:
\begin{align*}
    \Gamma_0 &= \varnothing^{\#} \setminus^{\#} \varnothing^{\#} \cup^{\#} \{\{A(0)\}\} = \{\{A(0)\}\}\\
    \Gamma_1 &= \{\{A(0)\}\} \setminus^{\#} \{\{A(0)\}\} \cup^{\#} \{\{B(0)\}\} = \{\{B(0)\}\}\\
    \Gamma_2 &= \{\{B(0)\}\} \setminus^{\#} \varnothing^{\#} \cup^{\#} \{\{A(0)\}\} = \{\{A(0), B(0)\}\}\\
    \Gamma_3 &= \{\{A(0), B(0)\}\} \setminus^{\#} \{\{B(0)\}\} \cup^{\#} \varnothing^{\#} = \{\{A(0)\}\}
\end{align*}
while the generated trace would be
\begin{equation*}
    tr = \langle \{\{\textrm{Init}(0)\}\}, \{\{\varnothing^{\#}\}\}, \{\{\textrm{Init}(0)\}\}, \{\{\textrm{Concl}(0)\}\},\rangle
\end{equation*}
and the observable trace
\begin{equation*}
    tr_{obs} = \langle \{\{\textrm{Init}(0)\}\}, \{\{\textrm{Init}(0)\}\}, \{\{\textrm{Concl}(0)\}\}\rangle
\end{equation*}

@@

\begin{itemize}
    \item $[ \ ] \to [ \mono{Fr}(~msg) ]$ allows for the generation of new fresh values.
    \item $[ \mono{Fr}(~msg) ] \to [ \mono{K}(~msg) ]$ allows for the generation of new fresh values by the attacker (expressed by the $\mono{K}$ fact)
    \item $[ \mono{Out}(msg) ] \to [ \mono{K}(msg) ]$ allows the attacker to eavesdrop all messages travelling through the network. 
    \item $[ \mono{K}(msg) ] \xrightarrow{\mono{K}(msg)} [ \mono{In}(msg) ]$ allows user to retrieve messages from the attacker-controlled network.
    \item $[ \ ] \to [ \mono{K}(\$x) ]$ allows the attacker to discover all public names.
    \item $[ \mono{K}(x1,...,xn) ] \to [ \mono{K}(\mono{f}(x1,...,xn)) ]$ allows the attacker to apply $n$-ary functions to arguments he already knows.
\end{itemize}

@@

\begin{gather*}
    [ \mono{Out}_{\textrm{conf}}(msg)] \to [ \mono{In}_{\textrm{conf}}(x) ]\\
    [ \mono{K}(msg)]
\end{gather*}

>>>>

Given a multiset $\Gamma_t = \{ F_0, ..., F_n \}$ and a sequence of multisets $trace_t = \langle a_0, ..., a_{t-1} \rangle$ at a time $t$, we can define a rewrite rule as a triple of multisets $RR = \langle L, A, R \rangle$ (written as $RR= L \xrightarrow{ A } R$) such that:

\begin{itemize}
    \item we can apply $RR$ to $\Gamma_t$ if there is at least one ground instance (i.e. an instance with no variables) $rr = l \xrightarrow{ a } r$ of $RR$ so that $l \subseteq^{\#} \Gamma_t$
    \item applying $rr$ to $\Gamma_t$ yields to a new state $\Gamma_{t+1}$ and an increased trace $trace_{t+1}$ obtained as
\begin{align*}
    \Gamma_{t+1} &= \Gamma_t \setminus^{\#} lin(l) \cup^{\#} r\\
    trace_{t+1} &= \langle a_0, ..., a_{t-1}, a \rangle
\end{align*}
\end{itemize}
in which $\setminus^{\#}$ and $\cup^{\#}$ are the multiset equivalent operations for set difference and union and $lin(l)$ is the multiset of linear facts belonging to $l$ (notice that persistent facts are never removed from the state). From now on, we will refer to $L$, $R$ and $A$ as the multisets of \textit{premises}, \textit{conclusions} and \textit{action facts} of a rule (in this order).

@@

\begin{definition}[Set of possible traces]
    G Given a set of labelled rewriting rules $P = \{(N_1,RR_1),$ $...,(N_m,RR_m)\}$, we define the set of possible traces generated by $P$ as
    \begin{align*}
        traces(P) = \{ &\langle A_1,...,A_n \rangle | \exists S_1,...,S_n . \varnothing ^{\#} \xrightarrow[]{A_1} S_1 \xrightarrow[]{A_2} ... \xrightarrow[]{A_n} S_n\\
        & \textrm{and no ground istance of } \mono{Fresh}() \textrm{is used twice }\}
    \end{align*}
    where $A_i$ is $RR_i$'s action facts multiset.
\end{definition}

@@

To better understand how such trace (both observable and not) is derived from a protocol's execution, we can take a look at the following example rewriting system:
\begin{align*}
    P = \{ &(N_1, [ \ ] \xrightarrow{\mono{Init}(0)} \{\{A(0)\}\}),\\
    &(N_2, \{\{A(x)\}\} \to \{\{B(x)\}\})\\
    &(N_3, \{\{B(x)\}\} \xrightarrow{\mono{Concl}(x)} [ \ ])\}
\end{align*}

Let us assume we are starting with an empty state $\Gamma_0$ and apply rules $N_1, N_2, N_1, N_3$, in this order. The state would evolve as:
\begin{align*}
    \Gamma_0 &= \varnothing^{\#} \setminus^{\#} \varnothing^{\#} \cup^{\#} \{\{A(0)\}\} = \{\{A(0)\}\}\\
    \Gamma_1 &= \{\{A(0)\}\} \setminus^{\#} \{\{A(0)\}\} \cup^{\#} \{\{B(0)\}\} = \{\{B(0)\}\}\\
    \Gamma_2 &= \{\{B(0)\}\} \setminus^{\#} \varnothing^{\#} \cup^{\#} \{\{A(0)\}\} = \{\{A(0), B(0)\}\}\\
    \Gamma_3 &= \{\{A(0), B(0)\}\} \setminus^{\#} \{\{B(0)\}\} \cup^{\#} \varnothing^{\#} = \{\{A(0)\}\}
\end{align*}
while the generated trace would be
\begin{equation*}
    tr = \langle \{\{\mono{Init}(0)\}\}, \{\{\varnothing^{\#}\}\}, \{\{\mono{Init}(0)\}\}, \{\{\mono{Concl}(0)\}\},\rangle
\end{equation*}
and the observable trace
\begin{equation*}
    tr_{obs} = \langle \{\{\mono{Init}(0)\}\}, \{\{\mono{Init}(0)\}\}, \{\{\mono{Concl}(0)\}\}\rangle
\end{equation*}

@@

\begin{itemize}
    \item $[ \ ] \to [ \mono{Fr}(~msg) ]$ allows for the generation of new fresh values.
    \item $[ \mono{Fr}(~msg) ] \to [ \mono{K}(~msg) ]$ allows for the generation of new fresh values by the attacker (expressed by the $\mono{K}$ fact)
    \item $[ \mono{Out}_{\textrm{ins}}(msg) ] \to [ \mono{K}(msg) ]$ allows the attacker to eavesdrop all messages travelling through the network. 
    \item $[ \mono{K}(msg) ] \xrightarrow{\mono{K}(msg)} [ \mono{In}_{\textrm{ins}}(msg) ]$ allows user to retrieve messages from the attacker-controlled network.
    \item $[ \ ] \to [ \mono{K}(\$x) ]$ allows the attacker to discover all public names.
    \item $[ \mono{K}(x_1,...,x_n) ] \to [ \mono{K}(\mono{f}(x_1,...,x_n)) ]$ allows the attacker to apply $n$-ary functions to arguments he already knows.
\end{itemize}

@@

\begin{gather*}
    [ \mono{Out}_{\textrm{conf}}(msg)] \to [ \mono{In}_{\textrm{conf}}(x) ]\\
    [ \mono{K}(msg) ] \to [ \mono{In}_{\textrm{conf}}(x) ]
\end{gather*}
